Sequential algorithms to split and merge terabyte-scale 3D images

### 2017-06-21 paper.tex
Updates about disks

This paper makes the following contributions:
\begin{itemize}
  \item We propose a set of algorithms to split and merge 3D images
    that do not fit in memory.
  \item We model the performance (execution time) of these algorithms,
    based on a simple but realistic disk I/O model.
  \item We evaluate our model and algorithms on a representative
    use-case involving a 1~TB image a hard drive and a solid-state
    drive.
  \item We compare the algorithms in a variety of conditions based on the validated model.
\end{itemize}
Section~\ref{sec:algos} presents our algorithms and models,
Section~\ref{sec:results} details the results and
Sections~\ref{sec:discussion} discusses them.

\section{Algorithms and Models}
\label{sec:algos}

@@

The following properties are straightforward applications of these
definitions:
\begin{itemize}
\item number of bytes per chunk: $\frac{Rb}{n}$
\item number of chunks that can fit in memory: $\frac{m.n}{Rb}$
\end{itemize}
For the sake of mathematical simplicity, the models below assume that
the reconstructed image and blocks are cubic. Under this assumption,
we have the following definitions and properties:
\begin{itemize}
\item Number of slices in a block: $\sqrt[3]{\frac{R}{n}}=d$. This is
  also the number of rows and columns in a slice and the number of
  voxel in a row or column.
\item Number of blocks required to fill a row of the reconstructed
  image: $\sqrt[3]{n}=\nu$.
\end{itemize}
\todo{Make a figure to illustrate the notations.}

@@

\subsection{Benchmark}
\label{sec:benchmark}

\subsubsection{Data}
As a running example to benchmark our algorithms, we use the 76~GB
BigBrain image split in 125 chunks. The data is available
\todo{[here]}.

\subsubsection{Infrastructure}
 We
perform experiments using the two disks below:
- HD ...
- SSD ...

\todo{Benchmark the disks to measure $\alpha_r$, $\alpha_w$ and $\beta$.}

\subsubsection{Software implementation}

Our algorithms are implemented in Python using the Nibabel
library. \todo{Software release.}

>>>>

This paper makes the following contributions:
\begin{itemize}
  \item We propose a set of algorithms to split and merge 3D images
    that do not fit in memory.
  \item We model the performance (execution time) of these algorithms,
    based on a simple but realistic disk I/O model.
  \item We evaluate our model and algorithms on a representative
    use-case involving a 1~TB image a hard drive and a solid-state
    drive.
  \item We compare the algorithms in a variety of conditions based on the validated model.
\end{itemize}
Section~\ref{sec:algos} presents our algorithms and their complexity,
Section~\ref{sec:results} details the results and
Sections~\ref{sec:discussion} discusses them.

\section{Algorithms}
\label{sec:algos}

@@

The following properties are straightforward applications of these
definitions:
\begin{itemize}
\item number of bytes per chunk: $\frac{Rb}{n}$
\item number of chunks that can fit in memory: $\frac{m.n}{Rb}$
\end{itemize}
For the sake of mathematical simplicity, we assume that
the reconstructed image and blocks are cubic. Under this assumption,
we have the following definitions and properties:
\begin{itemize}
\item Number of slices in a block: $\sqrt[3]{\frac{R}{n}}=d$. This is
  also the number of rows and columns in a slice and the number of
  voxel in a row or column.
\item Number of blocks required to fill a row of the reconstructed
  image: $\sqrt[3]{n}=\nu$.
\end{itemize}
\todo{Make a figure to illustrate the notations.}

@@

\section{Benchmark}
\label{sec:benchmark}

\subsubsection{Implementation}

Our algorithms are implemented in Python using the Nibabel
library. \todo{Val\'erie and Yongping: (1) extract ImageUtils from
  mapreduce repository (preserve git history; just copy the repo and
  remove the mapreduce stuff in the copy). Link to the github
  repo. (2) Describe the code briefly, in particular the
  implementation tricks used for each implementation.}

\subsubsection{Data}
To benchmark our library we use the 76~GB BigBrain image split in 125
non-overlapping chunks. The data is available \todo{[here]}.

\subsubsection{Infrastructure}

We run experiments using a Dell Precision Tower 3620 with CentOS Linux release
7.3.1611, 32~GB of RAM and two disks:
\begin{itemize}
  \item Hard disk drive (HDD): HGST Travelstar 7K1000, 7200~rpm, 931GiB (1TB), firmware version JB0OA3W0.
    \item Solid-state drive (SSD): SanDisk X400 2.5, 238GiB (256GB),
      firmware version X4130012.
\end{itemize}
Both drives used 512-byte logical sectors, 4096-byte physical sectors,
SATA >3.1 (6.0 Gb/s) and were accessed through the XFS file system.
%smartctl --xall /dev/sda
